---
title: 'DATA 607: Week 10'
author: "Cindy Lin"
date: "2025-04-01"
output:
  pdf_document: default
 
  
---

# INTRODUCTION

Week 10’s assignment focuses on text mining, and I chose to analyze poetry because it often relies less on individual words and more on overarching themes. Poetry tends to be highly nuanced, with meaning shaped significantly by the reader’s interpretation, making it an interesting challenge for text mining techniques. I selected Emily Dickinson’s work, as her poetry is frequently noted for its ambiguity and contrasting tones. I'm curious to see how sentiment analysis interprets her writing and whether the results align with the commonly held view of her work as layered and open to multiple interpretations.

I am starting by recreating some of the codes in the Text Mining textbook - 

**Citation:**
Silge, J., & Robinson, D. (2017). *Text Mining with R: A Tidy Approach*. O’Reilly Media. https://www.tidytextmining.com/sentiment.html

## Loading library

```{r library, include= FALSE}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
install.packages("tidytext", repos = "http://cran.us.r-project.org")
install.packages("gutenbergr", repos = "http://cran.us.r-project.org")
install.packages("jsonlite", repos = "http://cran.us.r-project.org")
install.packages("httr", repos = "http://cran.us.r-project.org")
install.packages("kableExtra", repos = "http://cran.us.r-project.org")



```
Installing the necessary packages for text mining - gutenbergr package to get Emily Dickinson's work. 

## Recreating with Emily Dickinson's WorK

```{r get Emily Dickinson work}

library(gutenbergr)
library(tidytext)
library(dplyr)
library(wordcloud)
library(sentimentr)
library(stringr)
library(knitr)


gutenberg_works(author == "Dickinson, Emily")

emily_poems <- gutenberg_download(12242)

tidy_emily <- emily_poems %>%
  unnest_tokens(word, text)
# make one word per row 

bing_word_counts <- tidy_emily %>%
  inner_join(get_sentiments("bing")) %>%
  count(sentiment, sort = TRUE) %>%
  ungroup()
# taken from the text 

print(bing_word_counts)


tidy_emily %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 50))

# word cloud
```
I searched Emily Dickinson's work in the Project Gutenberg collection and downloaded the full series which is the ID: 12242. By joining the token list (one word per row) and matching it to the bing sentiments, it will count up the positive and negative words. Based on the count, there are about the same positive words as negative words in the Emily Dickinson's series of poetry. With words like life, day, night, sun, death, thee, soul, time, heaven, god and mine being the most common words.


## Incorporate One Addiitional Sentiment
```{r}



dickinson <- gutenberg_download(12242)


text_combined <- paste(dickinson$text, collapse = " ")
#make into one long text


sentences <- get_sentences(text_combined)
#split into sentences using punctuation rules

flat_sentences <- unlist(sentences)
#make into vector

sentiment_scores <- sentiment(flat_sentences)

results <- cbind(sentiment_scores, sentence = flat_sentences)
#combine the score and sentences
overall_sentiment <- mean(results$sentiment)

if (overall_sentiment > 0) {
  cat("Overall sentiment is positive.\n")
} else if (overall_sentiment < 0) {
  cat("Overall sentiment is negative.\n")
} else {
  cat("Overall sentiment is neutral.\n")
}

# Get the overall sentiment score

print(overall_sentiment)

top_pos <- results[order(results$sentiment, decreasing = TRUE), ][1:5, ]
top_neg <- results[order(results$sentiment, decreasing = FALSE), ][1:5, ]




kable(top_pos, format = "latex")


kable(top_neg, format = "latex")




```
I wanted to use Sentimentr because it is a sentence level analysis which would be more appropriate in things like literature and poetry. The overall sentiment score of Emily Dickinson's work is neutral as the score is very close to zero. Because it was netural, I also wanted to do the top 5 postive and negative scores so we can see a general sense of her work from both end of the spectrum. 

## Using Azure Language Services

```{r}
library(httr)
library(jsonlite)

headers <- add_headers(
  `Content-Type` = "application/json",
  `Ocp-Apim-Subscription-Key` = Sys.getenv("api_key")

)

data <- '{"documents": [{"id": "2", "text": "
If I can stop one heart from breaking,
I shall not live in vain;
If I can ease one life the aching,
Or cool one pain,
Or help one fainting robin
Unto his nest again,
I shall not live in vain."}]}'

res <- POST(
  url = "https://week10data607.cognitiveservices.azure.com/text/analytics/v3.1/sentiment?opinionMining=false",
  headers,
  body = data,
  encode = "raw"
)

res
content(res)$documents[[1]]$sentiment


```
I am using the Azure Language Services and used a poem that I would imagine to be ambiguous in the individual words but the theme of the poem is inspirational and about compassion. This aligns with my thinking as the sentiment score is mixed. 

# CONCLUSION
By recreating the word analysis in Text Mining with R, we can better understand the differences in sentiment across various works. This comparison is particularly interesting because while the example in the text focused on literature, the text used here is poetry. Literature often invites a "read between the lines" interpretation, whereas poetry tends to be even more elusive and layered in meaning. By using sentimentr, we went beyond individual word sentiment analysis and explored the overall tone of the text. As expected, the sentiment analysis returned a neutral score, highlighting the subtlety and complexity of poetic expression.


 

